From Parametricity to Conservation Laws, via Noether's Theorem

### 2013-07-12 mechanics/paper.tex
Finish Sec 3, at last

FIXME: the following group homomorphisms become type-level constants,
using the interpretation we defined at the start of this section.

$\mathit{exp} : \mathsf{T}(1) \to \mathsf{GL}(1)$

$\underline{c}\cdot- : \mathsf{Z} \to \mathsf{T}(1)$

$\mathit{ortho}_n : \mathsf{O}(n) \to \mathsf{GL}(n)$

$\mathit{scale}_n : \mathsf{GL}(1) \to \mathsf{GL}(n)$

\paragraph{The Groupoid of Cartesian Spaces} Finally, we introduce the
groupoid kind of cartesian spaces, which we will use for the
configuration spaces of the classical mechanical systems we describe
in our system. As we noted above, the collection of cartesian spaces
is a groupoid with a non-trivial collection of objects. We extend our
system with a kind of cartesian spaces:
\begin{displaymath}
  \kappa ::= \cdots \sepbar \mathsf{CartSp}
\end{displaymath}
with the following interpretation:
\begin{displaymath}
  \begin{array}{@{}l}
    \sem{\mathsf{CartSp}} = \\
    \quad (\mathbb{N}, \lambda m,n.~\{ f : \mathbb{R}^m \to \mathbb{R}^n \sepbar f\textrm{ a diffeomorphism}\}, \lambda n.~\mathrm{id}_{\mathbb{R}^n})
  \end{array}
\end{displaymath}
We have chosen to represent cartesian spaces of dimension $n$ just as
the natural number $n$. This gives us a canonical representation of
$n$-dimensional space for each $n$, avoids tricky problems with
isomorphic constructions of the same space.

>>>>

We also extend our type system with the following type-level
constants. Each of these is interpreted by the correspondingly named
homomorphism between the groups involved. For example,
$\mathit{scale}_n$ is interpreted by the homomorphism $\mathrm{GL}(1)
\to \mathrm{GL}(n)$ that takes a non-zero real number $s$ to the
matrix with $s$s along the diagonal and zeros elsewhere.
\begin{displaymath}
  \begin{array}{l@{\hspace{0.3em}}c@{\hspace{0.3em}}l}
    \mathit{exp} &:& \mathsf{T}(1) \to \mathsf{GL}(1) \\
    \mathit{ortho}_n &:& \mathsf{O}(n) \to \mathsf{GL}(n) \\
    \mathit{scale}_n &:& \mathsf{GL}(1) \to \mathsf{GL}(n) \\
    \underline{c}*- &:& \mathsf{Z} \to \mathsf{T}(1) 
  \end{array}
\end{displaymath}
In the last of these $\underline{c}$ stands for an arbitrary real
constant.  Since each of these is interpreted by a group homomorphism,
we are justified in adding the group homomorphism laws as axioms to
our type-level equational theory, maintaining
Property~\ref{property:semantic-types}.

\paragraph{The Groupoid of Cartesian Spaces} Finally in this section,
we introduce the groupoid kind of cartesian spaces, which we will use
for the configuration spaces of the classical mechanical systems we
describe in our system. As we noted above, the collection of cartesian
spaces is a groupoid with a non-trivial collection of objects. We
extend our system with a kind of cartesian spaces:
\begin{displaymath}
  \kappa ::= \cdots \sepbar \mathsf{CartSp}
\end{displaymath}
with the following interpretation:
\begin{displaymath}
  \begin{array}{@{}l}
    \sem{\mathsf{CartSp}} = \\
    \quad (\mathbb{N}, \lambda m,n.~\{ f : \mathbb{R}^m \to \mathbb{R}^n \sepbar f\textrm{ a diffeomorphism}\}, \lambda n.~\mathrm{id}_{\mathbb{R}^n})
  \end{array}
\end{displaymath}
We have chosen to represent cartesian spaces of dimension $n$ just as
the natural number $n$. This gives us a canonical representation of
$n$-dimensional space for each $n$, avoids tricky problems with
isomorphic constructions of the same space.
